---
title: "Intro to Data Science - Group1 Project1"
author: "Group 1"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

#Question1

```{r, echo=FALSE}
library(readr)
library(readxl)
library(utils)
Nasdaq <- read_csv("Nasdaq 100 Historical Data.csv")
SP500 <- read_csv("S&P 500 Historical Data.csv")
USDIndex <- read_csv("US Dollar Index Historical Data.csv") 
Nasdaq$Date <- as.Date(Nasdaq$Date, format = "%m/%d/%Y")
Nasdaq <- subset(Nasdaq, select = -Vol.)
SP500$Date <- as.Date(SP500$Date, format = "%m/%d/%Y")
SP500 <- subset(SP500, select = -Vol.)
USDIndex$Date <- as.Date(USDIndex$Date, format = "%m/%d/%Y")
USDIndex <- subset(USDIndex, select = -Vol.)
Nasdaq=Nasdaq[rev(1:nrow(Nasdaq)), ]
SP500=SP500[rev(1:nrow(SP500)), ]
USDIndex=USDIndex[rev(1:nrow(USDIndex)), ]
SP500$Price <- as.numeric(gsub(",", "", SP500$Price))
Nasdaq$Price <- as.numeric(gsub(",", "", Nasdaq$Price))

str(Nasdaq)
str(USDIndex)
str(SP500)
missing_values <- sum(is.na(Nasdaq))
missing_values <- sum(is.na(SP500))
```

#   1-    Histograms

```{r, echo=FALSE}
library(ggplot2)
ggplot(Nasdaq, aes(x = Price)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black") +
  labs(title = "Nasdaq Price Histogram", x = "Price", y = "Frequency")
ggplot(USDIndex, aes(x = Price)) +
  geom_histogram(binwidth = 0.5, fill = "green", color = "black") +
  labs(title = "USDIndex Price Histogram", x = "Price", y = "Frequency")
ggplot(SP500, aes(x = Price)) +
  geom_histogram(binwidth = 10, fill = "red", color = "black") +
  labs(title = "SP500 Price Histogram", x = "Price", y = "Frequency")
```

Each histogram shows us that the different variable: "US Dollar index", "Nasdaq Index" and "S&P500" are normally distributted with different skewedness acroos the plot.

# 2-    Boxplot for each variables

```{r, echo=FALSE}
library(ggplot2)
ggplot(Nasdaq, aes(y = Price)) +
  geom_boxplot() +
  labs(title = "Nasdaq Price Box Plot")
ggplot(USDIndex, aes(y = Price)) +
  geom_boxplot() +
  labs(title = "USDIndex Price Box Plot")
ggplot(SP500, aes(y = Price)) +
  geom_boxplot() +
  labs(title = "SP500 Price Box Plot")
```

Accoording to these boxplots, we only have outliers with our Nasdaq variable.

# 3-    Merge the data frames to filter USDIndex by Nasdaq's dates

```{r, echo=FALSE}
merged_data <- merge(Nasdaq, USDIndex, by = "Date")
str(merged_data)
```

# 4-    Correlation between Nasdaq & US Dollar price Index

```{r, echo=FALSE}
corr_coef <- cor(merged_data$Price.x, merged_data$Price.y)
cat("Pearson Correlation Test :", corr_coef, "\n")
```

A correlation coefficient of 0.71246 between the US Dollar Index and the Nasdaq Index Price suggests a relatively strong positive linear relationship between these two variables. Meaning that when the US Dollar Index increases, the Nasdaq Index tends to increase as well.

# 5-    Line Plots
#   a-    Nasdaq Index & US Dollar Index

```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)

plot_nasdaq <- ggplot(data = merged_data, aes(x = Date)) +
  geom_line(aes(y = Price.x, color = "Nasdaq")) +
  labs(x = "Date", y = "Price") +
  scale_color_manual(values = c("Nasdaq" = "blue")) +
  ggtitle("Nasdaq Index Price")
plot_USDollar <- ggplot(data = merged_data, aes(x = Date)) +
  geom_line(aes(y = Price.y, color = "US Dollar")) +
  labs(x = "Date", y = "Price") +
  scale_color_manual(values = c("US Dollar" = "red")) +
  ggtitle("US Dollar Index Price")
grid.arrange(plot_nasdaq, plot_USDollar, ncol = 2)
```

These two line plots are showing a similar trend. With each upward or downward trend noticed in Nasdaq plot, we can aslo find similar trend on US Dollar plot at any time t.

#   b-    Nasdaq and S&P 500

```{r, echo=FALSE} 
ggplot() +
  geom_line(data = Nasdaq, aes(x = Date, y = Price, color = "Nasdaq"), size = 0.75) +
  geom_line(data = SP500, aes(x = Date, y = Price, color = "S&P 500"), size = 0.75) +
  labs(
    title = "Nasdaq and S&P 500 Comparison",
    x = "Date",
    y = "Price",
    color = "Index"
  ) +
  scale_color_manual(values = c("Nasdaq" = "green", "S&P 500" = "lightblue")) +
  theme_minimal()
```

From this comparison line plot, we can infer that Nasdaq and S&P500 have a positive relationship. Even though subtle, we can follow the movement of both variables across the chosen time frame and note that in every increase in S&P500 Index, there is also a quite significant increase in Nasdq Index.

# 6-    Scatter plot

```{r, echo=FALSE}
ggplot(data = SP500, aes(x = Price, y = Nasdaq$Price)) +
     geom_point() +
     labs(x = "S&P 500 Price", y = "Nasdaq Price") +
     ggtitle("Scatter Plot of S&P 500 Price vs. Nasdaq Price")
```

With this scatter plot, we can better identify the positive relationship betwwen S$P 500 Index and Nasdaq Index. we can infer that with every increase in Nasdaq, we can expect S&P 500 to increase as well.
